---
title: "02_uganda_uk"
author: "Theda Bartolomaeus and Till Birkner"
date: "2/17/2022"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

Dataset from Uganda + UK cohort, processed and rarefied together (Lotus and RTK) Lotus pipline 2 -\> SVL138 load packages

```{r echo=TRUE, message=FALSE, results = FALSE}
# Both dataset have been preprocessed together 
library(DirichletMultinomial)
library(reshape2)
library(magrittr)
library(dplyr)
library(cowplot)
library(ggrepel)
library(gridExtra)
library(grid)
library(ggplot2)
library(vegan)
library(lattice)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(egg)
```

Community typing will be performed in the merged UK + Uganda dataset

```{r}
OTU_uuk <-
  read.table(
    here::here(
      "input",
      "Lotus2_Uganda_SLV138",
      "higherLvl",
      "Rarefaction",
      "OTU_rarefied_to_8278.000000_n_0.tsv"
    ),
    sep = "\t",
    row.names = 1,
    header = T
  )
OTU_uuk$Rarefied <- NULL
md_uk <-
  read.table(
    here::here("input", "Metadata_UK_HIV.csv"),
    sep = ",",
    header = T,
    row.names = 1
  )
md_com <- md_uk[c(24, 28)] # select only hiv_status and geo_loc
colnames(md_com)[which(names(md_com) == "geo_loc_name_country_continent")] <-
  "geo_loc"
colnames(md_com)[which(names(md_com) == "host_hiv_status")] <-
  "Hiv_test"
load(here::here("input", "pretty_metadata_revision.r"))
### load metadata of both studies
metadata_Uganda <- metadata_revison
metadata_Uganda$geo_loc <- "Uganda"
metadata_Uganda <- metadata_Uganda[c(12, 42)]
colnames(metadata_Uganda)[which(names(metadata_Uganda) == "hiv_status")] <-
  "Hiv_test"
# metadata_Uganda <- metadata_Uganda[c(12,40)]
### now use cbind to bind both metadataframe
metadata_comb <- rbind(metadata_Uganda, md_com)
metadata_comb$Hiv_test[metadata_comb$Hiv_test == "No"] <- 0
metadata_comb$Hiv_test[metadata_comb$Hiv_test == "Yes"] <- 1
rownames(metadata_comb) <- sub("\\-", ".", rownames(metadata_comb))

```

```{r}
genus_uuk <-
  read.table(
    here::here(
      "input",
      "Lotus2_Uganda_SLV138",
      "higherLvl",
      "Rarefaction",
      "Genus_rarefied_to_8278.000000_n_0.tsv"
    ),
    sep = "\t",
    row.names = NULL
  )

# Remove non bacteria
genus_uuk <-
  genus_uuk[grep(patter = "Bacteria", x = genus_uuk$row.names),]
genus_uuk$Rarefied <- NULL
rownames(genus_uuk) <-
  make.names(genus_uuk$row.names, unique = TRUE)
genus_uuk$row.names <- NULL
#remove names of genus
raw_v <- as.data.frame(t(genus_uuk))
## Remove columnsum = 0
raw_v <- raw_v[, colSums(raw_v) > 0]
raw_v_t <- as.data.frame(t(raw_v))
#Note that genus_table, columns are samples and rows are genus
genus_table_uuk <- raw_v_t
dim(genus_table_uuk)# subset the data, genus_lvl count table (no proportion table)
head(colnames(genus_table_uuk)) #colnames should be samples
head(row.names(genus_table_uuk)) #rownames should be genus
# use the stringsplit function to get proper genus names:
# Archaea.Halobacterota.Methanosarcinia.Methanosa..-> to short
genus_puk <- as.data.frame(row.names(genus_table_uuk))
colnames(genus_puk) <- "long_names"
pretty_names_uuk <- tidyr::separate(
  genus_puk,
  long_names,
  into = c("A", "B", "C", "D", "E", "F"),
  sep = "\\.",
  fill = "right",
  extra = "drop"
)
write.table(pretty_names_uuk,
            here::here("output", "genus_names_pretty_uuk.tsv"))

# add short names to genus_table
genus_table_uuk <-
  as.data.frame(genus_table_uuk) #genus_table was integr
genus_table_uuk$short <- pretty_names_uuk$F
row.names(genus_table_uuk) <-
  make.names(genus_table_uuk$short, unique = T) #overcome duplicate row.names are not allowed
genus_table_uuk$short <- NULL
genus_table_uuk = genus_table_uuk / min(genus_table_uuk[genus_table_uuk >
                                                          0])

# Fit dirichlet multinomial model
all_dmns = 6 #max dirichlets to check for minimum information criteria
if (!file.exists(here::here("output", "dmn_list_uganda_uk.tsv"))) {
  dmn_list = numeric(all_dmns)
  for (i in 1:all_dmns) {
    print(i)
    assign(paste0("dmn_", i), dmn(as.matrix(t(
      genus_table_uuk
    )), i, verbose = F))
  }
  dmn_list = list(dmn_1, dmn_2, dmn_3, dmn_4, dmn_5, dmn_6)
  save(dmn_list, file = here::here("output", "dmn_list_uganda_uk.tsv"))
} else {
  load(here::here("output", "dmn_list_uganda_uk.tsv"))
  print("loaded dmn_list from saved file!")
}

lplc <- sapply(dmn_list, laplace)
plot(lplc,
     type = "b",
     xlab = "Number of Dirichlet Components",
     ylab = "Model Fit")
# best fit ist 5 The best fit is for k = 5distinct Dirichletcomponents.
#save clusters / metacommunities
Dirichlet_multinomial_1 = mixture(dmn_list[[1]], assign = TRUE)
Dirichlet_multinomial_2 = mixture(dmn_list[[2]], assign = TRUE)
Dirichlet_multinomial_3 = mixture(dmn_list[[3]], assign = TRUE)
Dirichlet_multinomial_4 = mixture(dmn_list[[4]], assign = TRUE)
Dirichlet_multinomial_5 = mixture(dmn_list[[5]], assign = TRUE)
Dirichlet_multinomial_6 = mixture(dmn_list[[6]], assign = TRUE)

Dirichlet_multinomial_all = data.frame(
  cbind(
    Dirichlet_multinomial_1,
    Dirichlet_multinomial_2,
    Dirichlet_multinomial_3,
    Dirichlet_multinomial_4,
    Dirichlet_multinomial_5,
    Dirichlet_multinomial_6
  )
)
colnames(Dirichlet_multinomial_all) = c("DMM_k=1",
                                        "DMM_k=2",
                                        "DMM_k=3",
                                        "DMM_k=4",
                                        "DMM_k=5",
                                        "DMM_k=6")

# minimum information criteria
lplc <- sapply(dmn_list, laplace)
BIC <- sapply(dmn_list, BIC)
AIC <- sapply(dmn_list, AIC)
dmn_list[[which.min(lplc)]] # optimal number of metacommunities (Laplace information criterion)
dmn_list[[which.min(BIC)]] # optimal number of metacommunities (Bayesian information criterion)
dmn_list[[which.min(AIC)]]# optimal number of metacommunities (Akaike information criterion)

# plot information criteria
plot(
  lplc,
  type = "b",
  xlab = "Number of Dirichlet Components",
  ylab = "Model Fit, Laplace",
  main = "Model fit as a function of Dirichlet component number"
)
# k = 4
plot(
  BIC,
  type = "b",
  xlab = "Number of Dirichlet Components",
  ylab = "Model Fit, BIC",
  main = "Model fit as a function of Dirichlet component number"
)
# K = 2
plot(
  AIC,
  type = "b",
  xlab = "Number of Dirichlet Components",
  ylab = "Model Fit, AIC",
  main = "Model fit as a function of Dirichlet component number"
)
# k = 2
```

```{r}
fit <- dmn_list
#best <- fit[[which.min(unlist(lplc))]] # proper way to do this
best <-
  fit[[4]] # manually set t 4 clusters to be socistent between differet pre-processing pipelines
#### so we want to get the weights of influence on the models from the different genera:
## relative abundance = contribution to each pulmotype(?)
cluster_imp_u <- fitted(best)

p1 <-
  fitted(fit[[1]], scale = TRUE)# vergleich gegen das model wo alles in einem
# gefittet ist
p5 <- fitted(best, scale = TRUE)

meandiff <- colSums(abs(p5 - as.vector(p1)))
meandiff
x <- mixture(best)

### plot the pulmotypes
plot_list = list()
for (k in seq(ncol(fitted(best)))) {
  d <- melt(fitted(best))
  colnames(d) <- c("OTU", "cluster", "value")
  d <- subset(d, cluster == k) %>%
    # Arrange OTUs by assignment strength
    arrange(value) %>%
    mutate(OTU = factor(OTU, levels = unique(OTU))) %>%
    # Only show the most important drivers
    filter(abs(value) > quantile(abs(value), 0.8))
  
  gg_color_hue <- function(n) {
    hues = seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
  }
  cols = gg_color_hue(4)
  
  p <- ggplot (d [(length(d$value) - 10):length(d$value),],
               aes(x = OTU , y = value)) +
    xlab("") +
    ylab("contribution to cluster") +
    geom_bar(stat = "identity",
             fill = cols[k],
             colour = "black") +
    coord_flip() +
    theme_classic() +
    theme(
      axis.text.y = element_text(face = "italic"),
      axis.text = element_text(size = 13, face = 'bold'),
      axis.title = element_text(size = 15, face = 'bold')
    )
  
  
  plot_list[[k]] = p
  print(paste0("Cluster importance plot done for cluster", k))
}

plot <- plot_grid(plot_list[[1]],
                  plot_list[[2]],
                  plot_list[[3]],
                  plot_list[[4]],
                  nrow = 2,
                  align = "vh")

print(length(plot_list))
# plot <- egg::ggarrange(plot_list[[1]], plot_list[[2]],
#                    plot_list[[3]],plot_list[[4]], 
#                    nrow = 2)
plot
```


changes the plots so they fit the figure structure 
```{r}
# use genus_table from community typing
genus_community <- as.data.frame(t(genus_table_uuk))
# reduce the dataset

# TB changed 3033_05_13
cluster_result_uuk <- as.data.frame(Dirichlet_multinomial_4)
# end change
ad.index.keep <- which(colSums(genus_community)*100/(sum(genus_community)) > 0.01) 
genus_community <- genus_community[, ad.index.keep]
dim(genus_community)
order(row.names(genus_community))
genus_community <- as.data.frame(genus_community)
genus_community$Smpl <- row.names(genus_community)
genus_community <- genus_community[rownames(genus_community) %in% rownames(metadata_comb),]
cluster_abundance <- merge(genus_community, cluster_result_uuk, by = 0)
row.names(cluster_abundance) <- cluster_abundance$Row.names
cluster_abundance$Row.names <- NULL
cluster_abundance$Smpl <- NULL

# MEDIAN
cluster_abundance_1_u <-
  apply(cluster_abundance[cluster_abundance$Dirichlet_multinomial_4 == "1",-ncol(cluster_abundance)], 2, mean)
cluster_abundance_2_u <-
 apply(cluster_abundance[cluster_abundance$Dirichlet_multinomial_4 == "2",-ncol(cluster_abundance)], 2, mean)
cluster_abundance_3_u <-
 apply(cluster_abundance[cluster_abundance$Dirichlet_multinomial_4 == "3",-ncol(cluster_abundance)], 2, mean)
cluster_abundance_4_u <-
 apply(cluster_abundance[cluster_abundance$Dirichlet_multinomial_4 == "4",-ncol(cluster_abundance)], 2, mean)

# merging
cluster_merged_u <-
  rbind(cluster_abundance_1_u,
        cluster_abundance_2_u,
        cluster_abundance_3_u,
        cluster_abundance_4_u)


## Plot the Abundance 
cluster_rel_u <- cluster_merged_u
cluster_rel_u <- as.data.frame(t(cluster_merged_u))
cluster_rel_u <- apply(cluster_merged_u, 1, function(x) x/sum(x)) # rows 1 als prozente

## use cluster_imp 
clr_u <- cluster_rel_u
clr_import_u <- merge(clr_u, cluster_imp_u, by = 0)
row.names(clr_import_u) <- clr_import_u$Row.names
clr_import_u$Row.names <- NULL


# use the clr_import
ru1 <- clr_import_u [, c(1,5)]
ru1 <- ru1[order(ru1$V1, decreasing = TRUE)[1:11],]
ru1$Genus <- row.names(ru1)

ru1 %>%
  mutate(Genus = fct_reorder(Genus, cluster_abundance_1_u)) %>%
  ggplot (aes(x = Genus, y = V1, fill = "tomato2"))+
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = round(cluster_abundance_1_u, digits = 3)), position = "dodge") +
  theme_classic() +
  coord_flip()+
   theme(axis.text.x = element_text(),
         axis.title.y = element_blank(),
        axis.text.y = element_text(face = "italic"),
        legend.position = "none") +
  ylab("cluster importance") -> rcomu1

rcomu1
```

```{r}
ru2 <- clr_import_u [, c(2,6)]
ru2 <- ru2[order(ru2$V2, decreasing = TRUE)[1:11],]
ru2$Genus <- row.names(ru2)

ru2 %>%
  mutate(Genus = fct_reorder(Genus, cluster_abundance_2_u)) %>%
  ggplot (aes(x = Genus, y = V2))+
  geom_bar(position="dodge", stat = "identity", color = "black", fill ="chartreuse3") +
  geom_text(aes(label = round(cluster_abundance_2_u, digits = 3)), position = "dodge")+
  theme_classic() +
  coord_flip()+
   theme(axis.text.x = element_text(),
         axis.title.y = element_blank(),
        axis.text.y = element_text(face = "italic"),
        legend.position = "none") +
  ylab("cluster importance") -> rcomU2
rcomU2
```

```{r}
ru3 <- clr_import_u [, c(3,7)]
ru3 <- ru3[order(ru3$V3, decreasing = TRUE)[1:11],]
ru3$Genus <- row.names(ru3)

ru3 %>%
  mutate(Genus = fct_reorder(Genus, cluster_abundance_3_u)) %>%
  ggplot (aes(x = Genus, y = V3))+
  geom_bar(position="dodge", stat = "identity", color = "black", fill ="deepskyblue2") +
  geom_text(aes(label = round(cluster_abundance_3_u, digits = 3)), position = "dodge")+
  theme_classic() +
  coord_flip()+
   theme(axis.text.x = element_text(),
         axis.title.y = element_blank(),
        axis.text.y = element_text(face = "italic"),
        legend.position = "none") +
  ylab("cluster importance") -> rcomU3
rcomU3
```

```{r}
ru4 <- clr_import_u [, c(4,8)]
ru4 <- ru4[order(ru4$V4, decreasing = TRUE)[1:11],]
ru4$Genus <- row.names(ru4)

ru4%>%
  mutate(Genus = fct_reorder(Genus, cluster_abundance_4_u)) %>%
  ggplot (aes(x = Genus, y = V4))+
  geom_bar(position="dodge", stat = "identity", color = "black", fill ="purple") +
  geom_text(aes(label = round(cluster_abundance_4_u, digits = 3)), position = "dodge")+
  theme_classic() +
  coord_flip()+
   theme(axis.text.x = element_text(),
         axis.title.y = element_blank(),
        axis.text.y = element_text(face = "italic"),
        legend.position = "none") +
  ylab("cluster importance") -> rcomU4
rcomU4
```
```{r}
rcomcomu <- cowplot::plot_grid(rcomu1, rcomU2, rcomU3,  rcomU4)
rcomcomu
```

bray-curtis dissimilarity UK vs Uganda colored according to their community type

```{r}
ComColu <- c("chartreuse3", "deepskyblue2", "red3", "purple")

cluster_result_uuk <- as.data.frame(Dirichlet_multinomial_4)
#Plot the Beta diversity colored for the different Clusters

md_uuk_com <- merge(metadata_comb, cluster_result_uuk, by = 0)
row.names(md_uuk_com) <- md_uuk_com$Row.names
md_uuk_com$Row.names <- NULL
# change the colnames, since they got a little messy
md_uuk_com$Dirichlet_multinomial_4 <-
  as.character(md_uuk_com$Dirichlet_multinomial_4)
#do pcoa
beta_uuk <- vegdist (t(OTU_uuk), method = "bray", na.rm = T)
## now we change the NAs to 0, since there is no diffrence between the samples
beta_uuk[is.na(beta_uuk)] <- 0
pcoaE <- cmdscale (beta_uuk, k = 2)
pcoaE <- as.data.frame(pcoaE)
head(pcoaE)

md_uuk_com <- merge(md_uuk_com, pcoaE, by = 0)
row.names(md_uuk_com) <- md_uuk_com$Row.names
md_uuk_com$Row.names <- NULL
md_uuk_com <- as.data.frame(md_uuk_com)

# prepare the plot
centroids <-
  aggregate(cbind(V1, V2) ~ Dirichlet_multinomial_4, md_uuk_com, mean)
md_uuk_com %>%
  ggplot (aes (x = V1, y = V2, color = Dirichlet_multinomial_4)) +
  theme_classic () +
  geom_point (
    aes (color = Dirichlet_multinomial_4),
    size = 5,
    alpha = 0.8,
    fill = cols[k]
  ) +
  xlab ("PCo 1") + ylab ("PCo 2") +
  theme (
    axis.title.x = element_text (size = 13),
    axis.text.x = element_text (size = 13),
    axis.text.y = element_text (size = 13),
    axis.title.y = element_text (size = 13),
    legend.position = "bottom"
  ) +
  stat_ellipse(aes(color = Dirichlet_multinomial_4)) +
  geom_point(
    data = centroids,
    size = 5,
    shape = 16,
    color = "black"
  ) + # centroides hinzufügen
  geom_point(data = centroids,
             size = 4,
             shape = 16) -> Bcluster_uuk

# Add density plots
md_uuk_com %>%
  ggplot(aes(x = V1), fill = cols[k]) +
  geom_density(alpha = .5,
               aes(fill = Dirichlet_multinomial_4,
                   color = Dirichlet_multinomial_4)) +
  # scale_fill_manual(values= cols[k]) +
  # scale_color_manual(values=cols[k]) +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    #axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    #axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  ) -> xdensity
#xdensity

md_uuk_com %>%
  ggplot(aes(V2), fill = cols[k]) +
  geom_density(alpha = .5,
               aes(fill = Dirichlet_multinomial_4,
                   color = Dirichlet_multinomial_4)) +
  # scale_fill_manual(values=pal.Coll) +
  # scale_color_manual(values=pal.Coll) +
  theme_classic() +
  theme(
    #axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    #axis.line.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  ) +
  theme(legend.position = "none") +
  coord_flip() -> ydensity
#ydensity

# Create blank plot to use for combining beta + the two desity plots
blankPlot <- ggplot() + geom_blank(aes(1, 1)) +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )
# bulid the final plot
# titnew = "Lotus2 SLV138 human reads filter"
Bpul_uuk <-
  plot_grid(
    xdensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    blankPlot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    Bcluster_uuk + theme(legend.position = "none",
                         plot.margin = unit(c(0, 0, 0, 0), "cm")),
    ydensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    nrow = 2,
    rel_widths = c(4, 1.4),
    rel_heights = c(1.4, 4),
    align = "hv"
  )
Bpul_uuk
```

more bet-div plots for HIV and UK vs. Uganda Geo_loc

```{r}
# prepare the plot
pal.Coll_uuk <- c("#04a3bd", "#20235b")
centroids <- aggregate(cbind(V1, V2) ~ geo_loc, md_uuk_com, mean)
md_uuk_com %>%
  ggplot (aes (x = V1, y = V2, color = geo_loc)) +
  theme_classic () +
  scale_color_manual(values = pal.Coll_uuk) +
  geom_point (aes (color = geo_loc),
              size = 5, alpha = 0.8) +
  xlab ("PCo 1") + ylab ("PCo 2") +
  theme (
    axis.title.x = element_text (size = 13),
    axis.text.x = element_text (size = 13),
    axis.text.y = element_text (size = 13),
    axis.title.y = element_text (size = 13),
    legend.position = "bottom"
  ) +
  stat_ellipse(aes(color = geo_loc)) +
  geom_point(
    data = centroids,
    size = 5,
    shape = 16,
    color = "black"
  ) + # centroides hinzufügen
  geom_point(data = centroids,
             size = 4,
             shape = 16) -> Bcluster_uuk

# Add density plots
md_uuk_com %>%
  ggplot(aes(x = V1)) +
  geom_density(alpha = .5, aes(fill = geo_loc,
                               color = geo_loc)) +
  scale_fill_manual(values = pal.Coll_uuk) +
  scale_color_manual(values = pal.Coll_uuk) +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    #axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    #axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  ) -> xdensity
#xdensity

md_uuk_com %>%
  ggplot(aes(V2)) +
  geom_density(alpha = .5, aes(fill = geo_loc,
                               color = geo_loc)) +
  scale_fill_manual(values = pal.Coll_uuk) +
  scale_color_manual(values = pal.Coll_uuk) +
  theme_classic() +
  theme(
    #axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    #axis.line.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  ) +
  theme(legend.position = "none") +
  coord_flip() -> ydensity
#ydensity

# Create blank plot to use for combining beta + the two desity plots
blankPlot <- ggplot() + geom_blank(aes(1, 1)) +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )
# bulid the final plot
Pul_uuk <-
  plot_grid(
    xdensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    blankPlot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    Bcluster_uuk + theme(legend.position = "none",
                         plot.margin = unit(c(0, 0, 0, 0), "cm")),
    ydensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    nrow = 2,
    rel_widths = c(4, 1.4),
    rel_heights = c(1.4, 4),
    align = "hv"
  )
Pul_uuk

```

Permanova - geo_loc

```{r}
# PERMANOVA 
permanova_geo <-
  adonis (vegdist (t(OTU_uuk), method = "bray", na.rm = T) ~ as.factor(geo_loc), data = md_uuk_com)
permanova_geo
```

Hiv_status

```{r}
# prepare the plot
pal.Coll_hiv <- c("#f0be3d", "#247d3f")
md_uuk_com <- md_uuk_com %>%
  na.omit()
centroids <- aggregate(cbind(V1, V2) ~ Hiv_test , md_uuk_com, mean)

md_uuk_com$Hiv_test[md_uuk_com$Hiv_test == "No"] <- 0
md_uuk_com %>%
  ggplot (aes (x = V1, y = V2, color = Hiv_test)) +
  theme_classic () +
  scale_color_manual(values = pal.Coll_hiv) +
  geom_point (aes (color = Hiv_test),
              size = 5, alpha = 0.8) +
  xlab ("PCo 1") + ylab ("PCo 2") +
  theme (
    axis.title.x = element_text (size = 13),
    axis.text.x = element_text (size = 13),
    axis.text.y = element_text (size = 13),
    axis.title.y = element_text (size = 13),
    legend.position = "bottom"
  ) +
  stat_ellipse(aes(color = Hiv_test)) +
  geom_point(
    data = centroids,
    size = 5,
    shape = 16,
    color = "black"
  ) + # centroides hinzufügen
  geom_point(data = centroids,
             size = 4,
             shape = 16) -> Bcluster_uuk

# Add density plots
md_uuk_com %>%
  ggplot(aes(x = V1)) +
  geom_density(alpha = .5, aes(fill = Hiv_test,
                               color = Hiv_test)) +
  scale_fill_manual(values = pal.Coll_hiv) +
  scale_color_manual(values = pal.Coll_hiv) +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    #axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    #axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  ) -> xdensity
#xdensity

md_uuk_com %>%
  ggplot(aes(V2)) +
  geom_density(alpha = .5, aes(fill = Hiv_test,
                               color = Hiv_test)) +
  scale_fill_manual(values = pal.Coll_hiv) +
  scale_color_manual(values = pal.Coll_hiv) +
  theme_classic() +
  theme(
    #axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    #axis.line.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  ) +
  theme(legend.position = "none") +
  coord_flip() -> ydensity
#ydensity

# Create blank plot to use for combining beta + the two desity plots
blankPlot <- ggplot() + geom_blank(aes(1, 1)) +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )
# bulid the final plot
Hivpul_uuk <-
  plot_grid(
    xdensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    blankPlot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    Bcluster_uuk + theme(legend.position = "none",
                         plot.margin = unit(c(0, 0, 0, 0), "cm")),
    ydensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    nrow = 2,
    rel_widths = c(4, 1.4),
    rel_heights = c(1.4, 4),
    align = "hv"
  )
Hivpul_uuk
```

Permanova hiv_status

```{r}
# PERMANOVA 
# drop na
# md_uuk_com <- md_uuk_com[!is.na(md_uuk_com$Hiv_test), ]
# rownames(OTU_uuk) %in% rownames(md_uuk_com)
# permanova_hiv <- adonis (vegdist (t(OTU_uuk), method = "bray", na.rm = T) ~ as.factor(Hiv_test),data = md_uuk_com)
# permanova_hiv
```

chiq testing and heatmap (geo_locs - community types)

```{r}
# md_uuk_com$Dirichlet_multinomial_4[metadata_beta$Dirichlet_multinomial_4 == 1] <- "Community 1"
# md_uuk_com$Dirichlet_multinomial_4[metadata_beta$Dirichlet_multinomial_4 == 2] <- "Community 2"
# md_uuk_com$Dirichlet_multinomial_4[metadata_beta$Dirichlet_multinomial_4 == 3] <- "Community 3"
# md_uuk_com$Dirichlet_multinomial_4[metadata_beta$Dirichlet_multinomial_4 == 4] <- "Community 4"

md_uuk_fram <- md_uuk_com
md_uuk_fram$Dirichlet_multinomial_4[md_uuk_fram$Dirichlet_multinomial_4 == "1"] <- "Community 1"
md_uuk_fram$Dirichlet_multinomial_4[md_uuk_fram$Dirichlet_multinomial_4 == "2"] <- "Community 2"
md_uuk_fram$Dirichlet_multinomial_4[md_uuk_fram$Dirichlet_multinomial_4 == "3"] <- "Community 3"
md_uuk_fram$Dirichlet_multinomial_4[md_uuk_fram$Dirichlet_multinomial_4 == "4"] <- "Community 4"

aFrame_com_1u <- reshape2::melt(t(apply(table (md_uuk_fram$geo_loc, md_uuk_fram$Dirichlet_multinomial_4),
                         1, function(x) x/sum(x))))

heatmapcomu <- ggplot (aFrame_com_1u, aes (x = Var1, y = Var2)) +
  theme_classic () +
  theme (axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
         legend.position = "none", 
         axis.title = element_blank()) +
  geom_tile (aes (fill = value), colour = "black") +
  geom_text (aes (label = scales::percent(value, accuracy = 1))) +
  scale_fill_gradient(high = "#d7301f", low = "#fff7ec") 
# scale_color_brewer(palette = "PiYG") +
print(heatmapcomu) 
```

look at alpha div

```{r}
rtk_div_uuk <-
  read.table(
    here::here(
      "input",
      "Lotus2_Uganda_SLV138",
      "higherLvl",
      "Rarefaction",
      "OTU_median_alpha_diversity.tsv"
    ),
    header = T,
    row.names = 1,
    sep = "\t"
  )

colnames(rtk_div_uuk) <-
  c("richness",
    "shannon",
    "simpson",
    "inv.simpson",
    "chao1",
    "eveness")
rtk_div_uuk <- rtk_div_uuk[-c(1),]
rownames(rtk_div_uuk) <- sub("\\-", ".", rownames(rtk_div_uuk))

#row.names(md_uuk_com) <- md_uuk_com$Row.names
#md_uuk_com$Row.names <- NULL

md_uuk_com <- merge(md_uuk_com, rtk_div_uuk, by = 0)
row.names(md_uuk_com) <- md_uuk_com$Row.names
md_uuk_com$Row.names <- NULL
md_uuk_com$shannon <- as.numeric(md_uuk_com$shannon)
md_uuk_com %>%
  # dplyr::group_by(DualStatus)%>%
  wilcox_test(shannon ~ geo_loc) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance() %>%
  add_xy_position(x = "geo_loc") -> stats_test


md_uuk_com %>%
  #dplyr::group_by(Line)%>%
  ggplot(aes(x = geo_loc, y = shannon)) +
  geom_boxplot(color = "black",
               alpha = 0.5,
               outlier.shape = NA) +
  geom_point(
    position = position_jitter(0.2),
    size = 3,
    aes(
      fill = geo_loc,
      color = geo_loc,
      shape = Hiv_test
    )
  ) +
  scale_color_manual(values = pal.Coll_uuk) +
  #scale_fill_npg() +
  #scale_shape_manual(values = c(21, 24))+
  xlab("DualStatus") +
  ylab("Microbiome richness (Shannon Index)") +
  labs(tag = "A)", caption = get_pwc_label(stats_test)) +
  theme_bw() +
  theme(
    text = element_text(size = 16),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x.bottom = element_text(size = 16),
    legend.position = "right"
  ) +
  stat_pvalue_manual(
    stats_test,
    bracket.nudge.y = -2,
    step.increase = 0.05,
    hide.ns = T,
    tip.length = 0,
    label = "{p.adj} {p.adj.signif}"
  ) -> A
A
```
plot shannon div as color gradient in beta-div plot: is alpha-div main driver of beta-div

```{r}
# dummy up data
dat1 <- md_uuk_com[md_uuk_com$geo_loc == "Uganda", ]
dat2 <- md_uuk_com[md_uuk_com$geo_loc == "Europe", ]

# geom_point with two scales
ggplot() +
       geom_point(data=dat1, aes(x = V1, y = V2, color = geo_loc, fill = shannon), shape=21, size=4) +
       geom_point(data= dat2, aes(x = V1, y = V2, color = geo_loc, fill = shannon), shape=21, size=4) +
       scale_fill_gradient(low="black", high="white")+
  theme_classic()

# Add density plots
md_uuk_com %>%
  ggplot(aes(x = V1)) +
  geom_density(alpha = .5, aes(fill = Hiv_test,
                               color = Hiv_test)) +
  scale_fill_manual(values = pal.Coll_hiv) +
  scale_color_manual(values = pal.Coll_hiv) +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    #axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    #axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  ) -> xdensity
#xdensity

md_uuk_com %>%
  ggplot(aes(V2)) +
  geom_density(alpha = .5, aes(fill = Hiv_test,
                               color = Hiv_test)) +
  scale_fill_manual(values = pal.Coll_hiv) +
  scale_color_manual(values = pal.Coll_hiv) +
  theme_classic() +
  theme(
    #axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    #axis.line.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  ) +
  theme(legend.position = "none") +
  coord_flip() -> ydensity
#ydensity

# Create blank plot to use for combining beta + the two desity plots
blankPlot <- ggplot() + geom_blank(aes(1, 1)) +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )
# bulid the final plot
Bpul_uuk <-
  plot_grid(
    xdensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    blankPlot + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    Bcluster_uuk + theme(legend.position = "bottom",
                         plot.margin = unit(c(0, 0, 0, 0), "cm")),
    ydensity + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
    nrow = 2,
    rel_widths = c(4, 1.4),
    rel_heights = c(1.4, 4),
    align = "hv"
  )
Bpul_uuk


```

metadeconfR run on features == phylum and metadata == geo_loc + community type 
```{r}
phylum_uuk <- read.table(here::here("input", "Lotus2_Uganda_SLV138", "higherLvl","Rarefaction", "Phylum_rarefied_to_8278.000000_n_0.tsv"), header = T)
phylum_uuk$Rarefied <- make.names(phylum_uuk$Rarefied, unique = T)
row.names(phylum_uuk) <- phylum_uuk$Rarefied
phylum_uuk$Rarefied <- NULL

# remove not bacteria 
phylum_uuk <- phylum_uuk[grep(patter= "Bacteria",x = row.names(phylum_uuk)), ]

phylum_puk <- as.data.frame(row.names(phylum_uuk))
colnames(phylum_puk) <- "long_names"
phylum_names_uuk <- tidyr::separate(
  phylum_puk,
  long_names,
  into = c("A", "B"),
  sep = "\\.",
  fill = "right",
  extra = "drop"
)
write.table(phylum_names_uuk,
            here::here("output", "phylum_names_pretty_uuk.tsv"))


# prepare metadeconf.R

phylum_uuk <- as.data.frame(t(phylum_uuk)) 
dim(phylum_uuk)
ad.index.keep <- which(colSums(phylum_uuk)*100/(sum(phylum_uuk)) > 0.01)
phylum_uuk <- phylum_uuk[, ad.index.keep]
dim(phylum_uuk)


md_uuk_metd <- md_uuk_fram[,- c(4:5)]
md_uuk_metd$smpl <- row.names(md_uuk_metd)
md_uuk_metd <- fastDummies::dummy_cols(.data = md_uuk_metd, select_columns = "Dirichlet_multinomial_4")
md_uuk_metd$Dirichlet_multinomial_4 <- NULL
row.names(md_uuk_metd) <- md_uuk_metd$smpl
md_uuk_metd$smpl <- NULL
phylum_uuk <- phylum_uuk[order(rownames(phylum_uuk)), ]
md_uuk_metd <- md_uuk_metd[order(rownames(md_uuk_metd)), ]

md_uuk_metd$Hiv_test[md_uuk_metd$Hiv_test == "No"] <- 0
md_uuk_metd$Hiv_test <- as.numeric(md_uuk_metd$Hiv_test)
md_uuk_metd <- md_uuk_metd %>%
  na.omit()

md_uuk_metd <- read.table(here::here("intermediate", "md_uuk_metd.tsv"), sep = "\t", header = T, row.names = 1)

md_uuk_metd$geo_loc <- as.numeric(md_uuk_metd$geo_loc)
phylum_uuk <- phylum_uuk[rownames(phylum_uuk) %in% rownames(md_uuk_metd), ]

meta_output_woConta_phylumuuk <- metadeconfoundR::MetaDeconfound(featureMat = phylum_uuk , 
                                   metaMat = md_uuk_metd, nnodes = 4, 
                                   logfile = here::here("intermediate", "MetadeconfoundR_wo_conta_phylum_uuk.log"))

mphylum_uuk <- metadeconfoundR::BuildHeatmap(meta_output_woConta_phylumuuk, d_col = c("blue", "white", "red"), d_range = "full")
## changes the names to italic
mphylum_uuk_1 <- mphylum_uuk +
  theme(axis.text.y = element_text(face ="italic"))
```


loook at the most abundance species in the Uganda - UK dataset 
```{r}
md_uuk_only <- md_uuk_metd[, -c (3:6)]
md_uuk_only$geo_loc[md_uuk_only$geo_loc == 1] <- "Uganda"
md_uuk_only$geo_loc[md_uuk_only$geo_loc == 0] <- "Europe"
md_uuk_only <- fastDummies::dummy_cols(.data = md_uuk_only, select_columns = "geo_loc")
meta_outputwocon_phylumuuk <- metadeconfoundR::MetaDeconfound(featureMat = phylum_uuk , 
                                   metaMat = md_uuk_only, nnodes = 4, returnLong = T,
                                   logfile = here::here("intermediate", "MetadeconfoundR_wo_conta_phylum_uuk_onlyloc.log"))

locphylum_uuk <- metadeconfoundR::BuildHeatmap(meta_outputwocon_phylumuuk, d_col = c("blue", "white", "red"), d_range = "full")
## changes the names to italic
locphylum_uuk_1 <- locphylum_uuk +
  theme(axis.text.y = element_text(face ="italic"))
locphylum_uuk_1
```




